---
title: "MS2Soft Functions Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MS2Soft Functions Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(candmR)
```

# Introduction

Read the MS2Soft Webscraping Vignette before working with the `ms2_` functions.

The `MS2Soft` group of functions within the `candmR` package are designed to
gather traffic data from the MS2 available platforms by US State. 
The `platforms` dataset contains the current state of available platforms, to add
a new platform to the package, communicate to the author. To add a temporary 
platform, use the input_file.xlsx file.

```{r platforms}

platforms

```

# Input Data

The `attributes` dataset contains an example of an attributes list which can be 
imported by using an inputs file. An example inputs file is included with the
package in the external data directory. A simplified version can be created with
the `ms2_inputs` function, which will create an MS Excel file with the corresponding
inputs and export it into the working directory.
Similarly, the `stations` dataset contains an example of how stations should be
organized before starting any queries. The `ms2_stationsCols` function creates
an MS Excel file containing the required columns to perform the analysis. An 
explanation of the contents of each column can be found in the package 
documentation.

```{r ms2_inputs, eval = FALSE}

ms2_inputs(path)
ms2_stationsCols(path)

```

Once an attributes file has been created, it can be imported as a list with the 
`ms2_attributes` function, which takes a single argument containing the path to
the inputs_file.xlsx file. This step can be skipped by more advanced users if
all attributes are defined within R.

```{r ms2_attrs, eval = FALSE}

attributes <- ms2_attributes(path)

```

The attributes list has 7 elements, which are explained in detail in the
package documentation. The following chunk shows the contents of the `attributes`
dataset included as an example.

```{r print_ms2_attrs}

str(attributes)

```

In the same way, the stations list file can be imported as a dataframe using the
`ms2_readstations` function. Which depends on having a stations Excel file like
with the configuration shown in the example data, or the one created with the
`ms2_stationsCols`.

```{r ms2_stns, eval = FALSE}

stations_list <- ms2_readStations(path)

```

The structure of the `stations` data is shown next, as with the `attributes`, a
detailed description of each variable can be found in the package documentation.

```{r print_ms2_stns}

str(stations)

```

# Data Query

The data scraping is separated into two phases, one where users download the
available dates, and one where users download the actual traffic data.
To perform any analysis, users must gather a list of available dates, this helps 
the program avoid errors due to bad requests, and also saves time, as
not all date links are visitted during the data download phase.

The `ms2_all_dates` function takes the stations list and the attributes object
as arguments and returns a list of all available dates within the specified
date span, which is defined per station within the `stations` data (columns 
`start_date` and `end_date`). The `ms2_all_dates` function relies on the 
`ms2_dates` function, which gathers all available dates for only one station.

```{r get_dates, eval = FALSE}

dates_list <- ms2_all_dates(stations_list = stations_list, attributes = attributes)

```

Next, use the `ms2_all_data` function to navigate to each of the URLs containing
the traffic data. Navigation is performed to reach each station, by given date 
(from the dates_list), by direction. And it's based on the analysis type defined
in the attributes. The function runs a Selenium session that navigates and 
downloads each file into the default downloads directory, and then copies them
into a "stations/" folder within the working directory. A subfolder referring to 
the analysis type is created, along with the corresponding years.

```{r get_data, eval = FALSE}

ms2_all_data(dates_list, stations_list = stations_list, attributes = attributes)

```

Finally, the `ms2_createdb` function takes the attributes object as an argument,
as well as a class number (for class type analyses) to determine the 
reclassiffication table to be used. This function gathers all downloaded files
for that specific analysis type and consolidates a database containing all
values.

```{r database, eval = FALSE}

ms2_createdb(attributes)

```

