---
title: "Google Maps Directions Usage Preparation and Prerequisites"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Google Maps Directions Usage Preparation and Prerequisites}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The Google Maps Platform allows to connect via an API and query travel times 
between two points. With the Directions API you can request routes for any
set of points and any date in the future, while also changing the time of the 
trip, so travel times at different times of day (TOD) can be consulted.
The `tt` group of function included in the **candmR** package allow to perform
a very specific method of travel time request, which is helpful to gather
times for trips along one or more corridor, separating by segments and direction.
The example shown here, takes one corridors on both directions that was segmented
at every major junction.

# Preparation

The **candmR** package is designed to have external files as inputs, given that it
relies on geographic data, which is easier to process within specialized GIS software
(although it can be created in R). The easiest and more straightforward way of 
defining the corridor routes, is to use Google Earth Pro, by setting the waypoints
directly in the map UI.

For this example, a segment of I-40 crossing Knoxville, Tennessee was cut down
into sub segments, at every major junction, for both directions of traffic.
The figure shows the corridor and the points that define the segments for each
route. For this example, the WB direction is named R1 and the EB direction is named
R2. Given the simplicity of the example, this is acceptable, but it would probably
make more sense to name each leg as R1_WB and R1_EB, or some other descriptive
code. Any convention works, the main recommendations would only be to keep names
short for ease of display, and to keep them consistent.


```{r segment, echo = FALSE, message = FALSE, out.width="80%"}

tmap::tmap_mode("view")
tmap::tm_shape(travel_times)+
  tmap::tm_lines(col = "route")+
  tmap::tm_shape(base_points)+
  tmap::tm_dots()

```


There are only three input datasets needed:
1. A dataframe containing the TOD attributes, one column with the TOD code and one
with the hour.
1. The points geographic file, preferrably as .kml generated in Google Earth
1. A points reference, where you define the correspondence of the points to the
routes.

The TOD attributes can be created with two methods, by defining a tod and an hours
vector or by feeding the TOD sheet from the inputs file. Both methods work under the
`tt_tod_data` function.

The points geographic data is reccommended to be created within Google Earth Pro,
by placing pins at the start at end of each segment. It is important to review that
the connection between the points returns the expected routing, by avoiding placing
the points outside of the road, or on an incorrect ramp or direction.

The PointDetails data, which should be generated within an MS Excel file should have
at least a Route and a Code column. This dataframe will be joined with the points
data using the `tt_gather_points` function. The new dataframe is one of the main
inputs of the 

```{r points, echo = FALSE, message = FALSE}

base_points

```

The actual process to gather travel times can be seen in the **travel times** 
vignette.
