---
title: "Google Maps Directions Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Google Maps Directions Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(candmR)
```

# Introduction

The `tt` family of functions were created to gather travel time data from
the Google Maps Directions API for a given road by using a set of 
waypoints along the chosen route. These waypoints determine the starting and
ending points and the required middle points that define the segments to be
queried. In that sense, spatial data for those points should be predefined
using either Google Earth or any other GIS software capable of producing a 
dataset in a format compatible with R (.kml and .shp have been tested).

Defining points is not covered in this document, but it should be noted that
points should be over the road segment and consider the direction of the 
traffic flow that will be downloaded. Testing the selected points within Google
Earth or Google Maps web platform is recommended to ensure that the route is
possible and only covers the required road segment.

# Defining Query Hours

There are two methods to define a `tod_list` (which is not a list object). The 
first method considers using the `tt_hours_vector` and defining a vector containing
the corresponding Time of Day (TOD). 

In the next chunk, a vector of tods is created which relates to a vector of hours,
created with the corresponding function. The `tt_hours_vector` requires a date
argument (which must be in the future), as well as a vector of hours in numeric 24
hour format. These hours relate to the tods vector created before.

```{r hours}
tods <- c("AM1", "MD1" ,"PM1", "PM1")
hours <- tt_hours_vector(Sys.Date()+1, c(8, 12, 16, 17))

```

The tod and the hours vectors can be used as an argument for the `tt_tod_data` 
function, which creates a tod_list, relating both datasets.

```{r tod_list}
tod_list <- tt_tod_data(tods, hours)

tod_list
```

The second option to import a tod list is to create an excel file with TOD and
Time columns and import it using the `tt_tod_data` function with the file path as
an attribute for the `table_path` argument.

```{r tod_path, eval = FALSE}
tod_list <- tt_tod_data(table_path = "path")

```

# Gather Points

The `tt_gather_points` function serves as an aid to create a special features object
containing important information about the waypoints to be requested, mainly the
Name (or coding) and the route correspondance. The function is designed to look
into two separate sources. The first is the geographic file containing the points,
which should have at least an ID parameter that can be related to the second source,
which should be an excel sheet containing the points attributes, mainly the order in
which the route should be followed and the route number, for multiple routes. In the
example route numbers are used to differentiate between the directions of the same
corridor.

```{r gather, eval = FALSE}

base_points <- tt_gather_points(points_path, point_details)

```

The result is a simple features object:

```{r gather_class}

class(base_points)

```

The previous function can be used, as long as the points object contains a geographic 
set of points, with an ID and a Route parameters and is sorted by the order in which
the route is followed, as shown here:

```{r print_points}

base_points

```

# Travel Time Query

The travel time query is performed using the `tt_get_traveltime` function. The
function requires the points data and the tod_list created in previous steps,
as well as the API key used to connect to the Google Maps platform. In addition,
the name and route parameters are helpers in case the Name and Route attributes
are called differently.

```{r times, eval = FALSE}

travel_times <- tt_get_traveltime(points, tod_list, key = key)

```

The travel_times data included in the package contains an example of one query for
two routes over 24 hours, with repeating TOD names.

# Summarizing data

The travel times functions of the candmR package include the `tt_matrix` and the
`tt_export` functions that help summarizing the obtained data, as well as 
creating working files with it, respectively.

The `tt_matrix` function can be used to summarize by *travel time*, *speed*
and *distance*. In addition, time of day periods can be grouped together in case
several hours are queried for the same time period. This grouping will take the
average value for the attribute consulted.
The most basic syntax for the `tt_matrix` function would take only the travel times
data obtained previously and will produce a list of matrices containing an 
element per route. In this case, a list with two elements, each containing a matrix
of times by route segment.

```{r matrix}

times <- tt_matrix(travel_times)
times

```

To add grouping, the argument group_tod should be set as TRUE:

```{r matrix_group}

times <- tt_matrix(travel_times, group_tod = TRUE)
times

```

To summarize another variable, use the summ_by attribute and set the parameter to 
either "speed" or "distance". The default value is "travel_times".

```{r matrix_speed}

speeds_tod <- tt_matrix(travel_times, summ_by = "speed", group_tod = TRUE)
speeds_tod

```

Finally, it is possible to export the results in three different formats:
A matrix format will create an excel file containing a sheet for each route,
a dataframe format will create an excel file containing the data from the 
travel_times object, while a shp format will create a geographic file, also
based on the travel_times object, containing all the routes consulted and the
data associated with them (for mapping).

```{r export, eval = FALSE}

tt_export(speeds_tod, "speedsTOD_matrix.xlsx", "matrix")
tt_export(travel_times, "travel_times.xlsx", "dataframe")
tt_export(travel_times, "travel_times.shp", "shp")

```
